\Sof{TBA - New Titles in the comments}

### 2020-04-30 paper/icsme20.tex
more updates

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of $10$ 
guidelines using a dataset including $1335$ security commits.
Results show that fixing software vulnerabilities may hinder the maintainability of 
open-source software. This is particularly the case for patches that 
address Path Traversal (CWE-22), Information Exposure (CWE-200) and Resource 
Management Errors (CWE-399). We conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

@@

From a methodological 
perspective, we leveraged a dataset of $1335$ security patches collected from open-source 
software. We calculate software maintainability before and after the patch to measure its 
impact. This empirical study presents evidence that changes 
applied in the code bases to patch vulnerabilities affect code maintainability.\Sof{fix}Especially, for patches that address Path Traversal (CWE-22),
Information Exposure (CWE-200) and Resource Management Errors (CWE-399). \Sof{-END-}This is important work
because there is still little information about the impact of security patches on the software
maintainability. With this study, 
we intend to highlight the need for tools to predict the impact of patches on 
maintainability~\cite{4724577}, documentation to help developers adopt security 
best practices~\cite{6311252, 7927935, MESQUIDA201519} and the importance of integrating
security-related code evolution in computer curricula.
%
%BCH compiles the ISO/IEC 25010 into a set of $10$ guidelines that are the
%base of BCH's evaluation. Two examples are the \emph{Write Simple Units of Code}
%based on the McCabe Complexity~\cite{1702388} and \emph{Keep your code base Small}
%based on the idea that smaller code bases are easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}.
%

@@

%%%%% NEEDS REVIEW
As an example, consider the patch of the TLS state-machine protocol implementation in 
OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting TLS messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2016$-$6307$~\footnote{CVE-$2016$-$6307$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2016-6307}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
research. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/statem/statem.c} file\footnote{CVE-$2016$-$6307$ fix available  at
\url{https://github.com/openssl/openssl/commit/c1ef7c971d0bbf117c3c80f65b5875e2e7b024b1}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. The problem in
CVE-$2016$-$6307$ is that due to a flaw in the logic of version $1.1.0$,
the memory for a message is allocated too early, prior to the excessive
message length check which can lead to a Denial of Service through memory
exhaustion if an attacker triggers up a certain message length to be allocated to service a
connection. In this case, there is only a security impact if the application does not call  \texttt{SSL\_free()} in a timely manner in the
event that the connection fails;
or, the application is working in a constrained environment where there
is very little free memory;
or, the attacker initiates multiple connection attempts such that there
are multiple connections in a state where memory has been allocated for
the connection, when \texttt{SSL\_free()} has not yet been called, and there is
insufficient memory to service the multiple requests. Except in the instance of $1$ above, any Denial Of Service is likely to
be transitory because as soon as the connection fails the memory is
subsequently freed again in the \texttt{SSL\_free()} call. However, there is an
increased risk during this period of application crashes due to the lack
of memory - which would then mean a more serious Denial of Service.

@@

The \texttt{read\_state\_machine} function, presented in Listing \ref{lst:vuln}, abstracts a sub-state-machine
to handle the messages exchanged between peers. When messages are larger than
what is defined by default for TLS, it may trigger a memory corruption issue since the allocation
of this resource was being verified very early in the application's flow, using the \texttt{tls\_get\_message\_header} function presented in Listing ~\ref{lst:vuln2} (\ref{lst:func3} and \ref{lst:func4}).
These verifications of size allocation were substituted by the one presented in Listing~\ref{lst:vuln}(\ref{lst:func2}).

%
Although checking the maximum size of a message seems an elementary
problem to repair, the example shows that there was a significant amount of
changes performed in the code base which produced a negative impact in the
maintainability. The developer tries to patch the
code base by introducing a new branch point to the function \texttt{read\_state\_machine}~\ref{lst:func1} which adds complexity to the software
and disrupts
one of the guidelines proposed by the Software Improvement Group (SIG) for
building maintainable software~\cite{Visser:2016:OREILLY} $-$
the \emph{Write Simple Units of Code} guideline $-$ which resulted in a negative impact on the
OpenSSL maintainability.
\Sof{-- END --}
%%%%%%% 
Software maintainability is designed as the degree to which an application is understood, 
repaired or enhanced. In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability of their
software applications. This is important because software maintainability is 
approximately $75\%$ of the cost related to a project. To answer the following three 
research questions, we use two datasets of security patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064} 
to measure the impact of
security patches on the maintainability of open-source software. 
%

@@

$1031$ security patches. $912$ of these patches are identified with Common Weakness

@@

The pair of the random commit and its parent is accepted if the \emph{diff} 
size fits in the range size considered acceptable at the moment. The algorithm 
starts with a step of $1$. Every $10$ attempts to search for a change with the 
same \emph{diff} the step increases. Thus, if a change suffered $20$ deletions 
and $10$ additions, then, in the first $10$ attempts, the algorithm searches 
for a \emph{diff} with deletions between $19$ and $21$ and additions between 
$9$ and $11$. After the $10$ first attempts, the algorithm increase the step 
for $2$ and searches for a \emph{diff} with deletions between $18$ and $22$ 
and additions between $8$ and $10$. We originate the regular commits from the 
security commits to ensure that differences in maintainability are not a 
consequence of characteristics of different projects.

@@

\Sof{Needs improvement}SIG defines \emph{Units} as the smallest groups of code that can be maintained
and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and
constructors in Java). One of the guidelines with which the project does not
comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}):
\emph{Write Simple Units of Code}. BCH analyzes this guideline based on the
McCabe Complexity~\cite{1702388} to calculate the number of branch points of a
method. The bar at the bottom of the figure represents the top $30$ units that
violate the guideline, sorted by severity. The violation severities are
indicated using colors, and there is a legend to help to interpret them. The
green bar represents the number of compliant branch points per unit (\emph{at
most $5$}), i.e., the number of units are compliant with ISO
$25010$~\cite{iso:2011}. Yellow, orange and red bars represent units that do not
comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint
the compliance thresholds for each severity level. If the green mark is
somewhere in the green bar it means it is compliant with a low level of
severity.\Sof{--END--}

@@

Aiming to analyze the impact of security patches, we use BCH to compute the
maintainability of two different versions of the project (cf. Figure~\ref{fig:commit}):
\begin{itemize}
	\item $v_{s-1}$, the version containing the security flaw, i.e., before the
	patch (\emph{sha-p});
	\item $v_{s}$, the version free of the security flaw, i.e., after the
	patch (\emph{sha});
\end{itemize}

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
The compliance for a severity level
is calculated based on the number of lines of code that comply and not comply
with the guideline at a given severity level as explained in more
detail in previous work~\cite{cruz2019energyoriented}. In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability. 
% The compliance $C$ for a given severity
% level $l$ is derived by:
%
% \begin{equation}\label{eq:3}
%     C(l) = LOC_{compliant}(l) - w(l) * LOC_{\neg compliant}(l)
% \end{equation}
%
% \noindent
% where $LOC_{compliant}(l)$ are the lines of code that comply with the guideline
% at the given severity level $l$, $LOC_{\neg compliant}(l)$ are the lines of code
% that do not comply with the guideline at the given severity level $l$ and $w(l)$
% is the weight factor to heighten the impact of non-compliant lines in comparison to
% compliant lines. Finally, the term $w(l)$ is calculated as follows:
%
% \begin{equation}
%     w(l) = \frac{1 - \theta(l)}{\theta(l)}
% \end{equation}
%
% \noindent
% where $\theta(l)$ is the threshold in percentage of the lines of code that are
% accepted to be non-compliant with the guideline for the severity level $l$. This
% is a standard value defined by BCH. In other words, the factor $w$ is used in
% Equation~\ref{eq:3} to highlight the lines of code that are not complying with
% the guideline. Then, we compute the difference of maintainability between the
% security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in
% Figure~\ref{fig:commit}.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\\\textit{\textbf{Construct:}} The maintainability formula was inferred based on the BCH's reports. The high
amount of different projects and backgrounds may require other
maintainability standards. However, BCH does use a representative benchmark of
closed and open-source software projects to compute the
thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
%
\\\textit{\textbf{Internal:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security patches that were
not explicit in commits messages. Pontas et. al\cite{10.1109/MSR.2019.00064} dataset
has patches that were manually mined in software repositories. Being a manual technique,
the search might led to
wrong associations between vulnerabilities and code snippets.
Baseline commits are retrieved randomly by selecting one commit from the same
project of a given security patch. This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $1031$ regular commits are enough to
alleviate random irregularities in the maintainability differences of the
baseline.
%
\\\textit{\textbf{External:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security patches in any other languages but English.

@@

This work presents an empirical study on the impact of $1031$ security
patches on the maintainability of $267$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability when patching vulnerabilities. We found this evidence in
$41.76\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability stays the same ($52.06\%$). However, there is still a 
considerable amount of cases that harm maintainability ($21.04\%$).

This study also shows evidence that there are security patterns that need more
attention than others. \emph{Path Traversal} (CWE-22),
\emph{Information exposure} (CWE-200) and \emph{Resource Management Errors} (CWE-399) vulnerabilities endure maintainability
in $48.6\%$, $51.2\%$, and $54.5\%$, respectively. Therefore, we draw attention
for the need of better documentation explaining how to use best practices to
produce secure and maintainable code; new and better programming languages; new
academic curricula in the fields of security and software quality; and, tools
such as BCH to help developers predict the effect of their patches.

>>>>

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of 10 
guidelines using a dataset including 1335 security commits.
Results show that fixing software vulnerabilities may hinder the maintainability of 
open-source software. This is particularly the case for patches that 
address Improper Input Validation (CWE-20), Information 
Exposure (CWE-200), Improper Restriction of Operations within the Bounds of a 
Memory Buffer (CWE-119), Missing Release of Memory after Effective Lifetime
(CWE-401) and Path Traversal (CWE-22). We conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

@@

From a methodological 
perspective, we leveraged a dataset of $1335$ security patches collected from open-source 
software. We calculate software maintainability before and after the patch to measure its 
impact. This empirical study presents evidence that changes 
applied in the code bases to patch vulnerabilities affect code maintainability. Especially, for patches that address \emph{Improper Input Validation} (CWE-20), \emph{Information 
Exposure} (CWE-200), \emph{Improper Restriction of Operations within the Bounds of a 
Memory Buffer} (CWE-119), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22). This is important work
because there is still little information about the impact of security patches on the software
maintainability. With this study, 
we intend to highlight the need for tools to predict the impact of patches on 
maintainability~\cite{4724577}, documentation to help developers adopt security 
best practices~\cite{6311252, 7927935, MESQUIDA201519} and the importance of integrating
security-related code evolution in computer curricula.
%
%BCH compiles the ISO/IEC 25010 into a set of $10$ guidelines that are the
%base of BCH's evaluation. Two examples are the \emph{Write Simple Units of Code}
%based on the McCabe Complexity~\cite{1702388} and \emph{Keep your code base Small}
%based on the idea that smaller code bases are easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}.
%

@@

%%%%% NEEDS REVIEW
\Sof{As an example, consider the patch of the TLS state-machine protocol implementation in 
OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting TLS messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2016$-$6307$~\footnote{CVE-$2016$-$6307$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2016-6307}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
research. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/statem/statem.c} file\footnote{CVE-$2016$-$6307$ fix available  at
\url{https://github.com/openssl/openssl/commit/c1ef7c971d0bbf117c3c80f65b5875e2e7b024b1}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. The problem in
CVE-$2016$-$6307$ is that due to a flaw in the logic of version $1.1.0$,
the memory for a message is allocated too early, prior to the excessive
message length check which can lead to a Denial of Service through memory
exhaustion if an attacker triggers up a certain message length to be allocated to service a
connection. In this case, there is only a security impact if the application does not call  \texttt{SSL\_free()} in a timely manner in the
event that the connection fails;
or, the application is working in a constrained environment where there
is very little free memory;
or, the attacker initiates multiple connection attempts such that there
are multiple connections in a state where memory has been allocated for
the connection, when \texttt{SSL\_free()} has not yet been called, and there is
insufficient memory to service the multiple requests. Except in the instance of $1$ above, any Denial Of Service is likely to
be transitory because as soon as the connection fails the memory is
subsequently freed again in the \texttt{SSL\_free()} call. However, there is an
increased risk during this period of application crashes due to the lack
of memory - which would then mean a more serious Denial of Service.}

@@

\Sof{The \texttt{read\_state\_machine} function, presented in Listing \ref{lst:vuln}, abstracts a sub-state-machine
to handle the messages exchanged between peers. When messages are larger than
what is defined by default for TLS, it may trigger a memory corruption issue since the allocation
of this resource was being verified very early in the application's flow, using the \texttt{tls\_get\_message\_header} function presented in Listing ~\ref{lst:vuln2} (\ref{lst:func3} and \ref{lst:func4}).
These verifications of size allocation were substituted by the one presented in Listing~\ref{lst:vuln}(\ref{lst:func2}).
%
Although checking the maximum size of a message seems an elementary
problem to repair, the example shows that there was a significant amount of
changes performed in the code base which produced a negative impact in the
maintainability. The developer tries to patch the
code base by introducing a new branch point to the function \texttt{read\_state\_machine}~\ref{lst:func1} which adds complexity to the software
and disrupts
one of the guidelines proposed by the Software Improvement Group (SIG) for
building maintainable software~\cite{Visser:2016:OREILLY} $-$
the \emph{Write Simple Units of Code} guideline $-$ which resulted in a negative impact on the
OpenSSL maintainability.}

%%%%%%% 
Software maintainability is designed as the degree to which an application is understood, 
repaired or enhanced. In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability of their
software applications. This is important because software maintainability is 
approximately $75\%$ of the cost related to a project. To answer the following three 
research questions, we use two datasets of security patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064} 
to measure the impact of
security patches on the maintainability of open-source software. 
%

@@

$998$ security patches. $912$ of these patches are identified with Common Weakness

@@

\Sof{The pair of the random commit and its parent is accepted if the \emph{diff} 
size fits in the range size considered acceptable at the moment. The algorithm 
starts with a step of $1$. Every $10$ attempts to search for a change with the 
same \emph{diff} the step increases. Thus, if a change suffered $20$ deletions 
and $10$ additions, then, in the first $10$ attempts, the algorithm searches 
for a \emph{diff} with deletions between $19$ and $21$ and additions between 
$9$ and $11$. After the $10$ first attempts, the algorithm increase the step 
for $2$ and searches for a \emph{diff} with deletions between $18$ and $22$ 
and additions between $8$ and $10$. We originate the regular commits from the 
security commits to ensure that differences in maintainability are not a 
consequence of characteristics of different projects.}

@@

\Sof{SIG defines \emph{Units} as the smallest groups of code that can be maintained
and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and
constructors in Java). One of the guidelines with which the project does not
comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}):
\emph{Write Simple Units of Code}. BCH analyzes this guideline based on the
McCabe Complexity~\cite{1702388} to calculate the number of branch points of a
method. The bar at the bottom of the figure represents the top $30$ units that
violate the guideline, sorted by severity. The violation severities are
indicated using colors, and there is a legend to help to interpret them. The
green bar represents the number of compliant branch points per unit (\emph{at
most $5$}), i.e., the number of units are compliant with ISO
$25010$~\cite{iso:2011}. Yellow, orange and red bars represent units that do not
comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint
the compliance thresholds for each severity level. If the green mark is
somewhere in the green bar it means it is compliant with a low level of
severity.}

@@

Aiming to analyze the impact of security patches, we use BCH to compute the
maintainability of two different versions of the project (cf. Figure~\ref{fig:commit}):

\begin{itemize}
	\item $v_{s-1}$, the version containing the security flaw, i.e., before the
	patch (\emph{sha-p});
	\item $v_{s}$, the version free of the security flaw, i.e., after the
	patch (\emph{sha});
\end{itemize}

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
\Sof{The compliance for a severity level
is calculated based on the number of lines of code that comply and not comply
with the guideline at a given severity level as explained in more
detail in previous work~\cite{cruz2019energyoriented}.} In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability. 
% The compliance $C$ for a given severity
% level $l$ is derived by:
%
% \begin{equation}\label{eq:3}
%     C(l) = LOC_{compliant}(l) - w(l) * LOC_{\neg compliant}(l)
% \end{equation}
%
% \noindent
% where $LOC_{compliant}(l)$ are the lines of code that comply with the guideline
% at the given severity level $l$, $LOC_{\neg compliant}(l)$ are the lines of code
% that do not comply with the guideline at the given severity level $l$ and $w(l)$
% is the weight factor to heighten the impact of non-compliant lines in comparison to
% compliant lines. Finally, the term $w(l)$ is calculated as follows:
%
% \begin{equation}
%     w(l) = \frac{1 - \theta(l)}{\theta(l)}
% \end{equation}
%
% \noindent
% where $\theta(l)$ is the threshold in percentage of the lines of code that are
% accepted to be non-compliant with the guideline for the severity level $l$. This
% is a standard value defined by BCH. In other words, the factor $w$ is used in
% Equation~\ref{eq:3} to highlight the lines of code that are not complying with
% the guideline. Then, we compute the difference of maintainability between the
% security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in
% Figure~\ref{fig:commit}.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\\\textit{\textbf{Construct:}} The maintainability formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
%
\\\textit{\textbf{Internal:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security patches that were
not explicit in commits messages. Pontas et. al\cite{10.1109/MSR.2019.00064} dataset
has patches that were manually mined in software repositories. Being a manual technique,
the search might led to
wrong associations between vulnerabilities and code snippets.
Baseline commits are retrieved randomly by selecting one commit from the same
project of a given security patch. This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $998$ regular commits are enough to
alleviate random irregularities in the maintainability differences of the
baseline.
%
\\\textit{\textbf{External:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security patches in any other languages but English.

@@

This work presents an empirical study on the impact of $998$ security
patches on the maintainability of $262$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018, cruz2019energyoriented}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability when patching vulnerabilities. We found this evidence in
$41.28\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability stays the same ($42.69\%$). However, there is still a 
considerable amount of cases that harm maintainability ($27.05\%$).

This study also shows evidence that there are security patterns that need more
attention than others. \emph{Improper Input Validation} (CWE-20), \emph{Information 
Exposure} (CWE-200), \emph{Improper Restriction of Operations within the Bounds of a 
Memory Buffer} (CWE-119), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22) endure maintainability
in $41.9\%$, $52.5\%$, $42.9\%$, $49.4\%$, and $41.7\%$, respectively. Therefore, 
we draw attention for the need of better documentation explaining how to use best practices to
produce secure and maintainable code; new and better programming languages; new
academic curricula in the fields of security and software quality; and, tools
such as BCH to help developers predict the effect of their patches.
